Which Scenario Should We Design For? Insights from House Elevation on the Multiple PDF Problem

### 2022-01-09 papers/paper/preprint.tex
Get the stuff from poster into preprint -- now need to outline and flesh out

% Figures
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{{../plots/}{../tikz/}{../img/}}
\usepackage[section]{placeins} % require floats to appear in the section they are defined

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, fleqn, nicefrac}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}

@@

@@

% use biblatex
\usepackage{csquotes}
\usepackage[
  backend=biber,
  doi=true,
  url=false,
  isbn=false,
  style=authoryear-comp,
  natbib=true,
  backref=false,
  maxbibnames=10,
  maxcitenames=2,
  uniquename=false,
  uniquelist=false,
  sorting=nyt,
]{biblatex}
\renewbibmacro{in:}{}
\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
%\addbibresource{../_bibliography/library.bib}
\addbibresource{library.bib}

@@

\section*{Front Matter}

\subsection*{Submission Plan}
We should revisit this!
\begin{enumerate}
    \item Earth's Future
    \item Climatic Change
    \item IOP Environmental Research: Infrastructure and Sustainability (new journal)
    \item Risk analysis
\end{enumerate}
\subsection*{Internal peer review}
\begin{enumerate}
    \item Vivek
    \item Courtney
    \item Sanjib
    \item Kelsey to replicate code
\end{enumerate}

@@

\begin{figure}
    \centering
    \includegraphics[height=2in]{uconn-elevateA}
    \caption{
        Conceptual sketch illustrating the path from historical data to FEMA's guidance.
        This figure \emph{needs to be revised} (KATERINA?) to have three panels, of which this is the third.
        (a): A time series of historical floods.
        (b): A \gls{pdf} of the flood peak.
        (c): as shown, but with waves eliminated.
        This sketch figure is from \citet{patten_elevate:2016}.
    }\label{fig:house-sketch}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth, height=4.5in, keepaspectratio=true]{conceptual.pdf}
    \caption{
        Schematic of our decision support framework.
        Nodes in green are modeling choices that the modeler must make while those in yellow are computed quantities.
        Teal text illustrates a non-exhaustive set of opportunities to perform model checks and/or engage with stakeholders to iterate on design of the decision support system.
        Feedbacks between these questions and the green nodes are omitted for clarity.
        Our approach could be extended by coupling the evaluation framework to an optimization or policy search tool that updates the candidate decisions.
        This approach could also be extended to a stochastic system model.
    }\label{fig:framework}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[height=2.5in, width=0.6\textwidth, keepaspectratio=true]{xlrm}
    \caption{
        Diagram showing uncertainties ($X$; red), levers ($L$; yellow), system relationships ($R$; gray), and metrics ($M$; blue).
    }\label{fig:xlrm}
\end{figure}

>>>>

% Figures
\usepackage{graphicx}
\usepackage[list=true]{subcaption}
\graphicspath{{../../plots/}{../../tikz/}{../../img/}}
\usepackage[section]{placeins} % require floats to appear in the section they are defined

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, fleqn, nicefrac}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}
\usepackage{fontawesome5}

@@

\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}

@@

% use biblatex
\usepackage{csquotes}
\usepackage[
  backend=biber,
  doi=true,
  url=false,
  isbn=false,
  style=authoryear-comp,
  natbib=true,
  backref=false,
  maxbibnames=10,
  maxcitenames=2,
  uniquename=false,
  uniquelist=false,
  sorting=nyt,
]{biblatex}
\renewbibmacro{in:}{}
\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
\addbibresource{library.bib}

@@

\clearpage

\section{Introduction}

Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \citep{zarekarizi_suboptimal:2020,xian_elevation:2017}.
\begin{description}
    \item[Q1] How to adapt guidance to building characteristics or household preferences?
    \item[Q2] How does nonstationary hazard change guidance?
    \item[Q3] Which model of nonstationary hazard to use?
\end{description}
These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).

The paper is organized as follows.
In \cref{sec:case-study} we describe our case study in Norfolk, VA and present results to illustrate the ``Multiple PDF problem'' and motivate our methodological advances.
Then in \cref{sec:prior-averaging}, we present a method for combining multiple \glspl{pdf}.
Finally in \cref{sec:conclusions} we summarize our findings, discuss implications for engineering practice, and identify needs for future research.

\section{Case study}\label{sec:case-study}

We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard.

\subsection{Model structure and data}

We follow the general modeling approach of \citet{zarekarizi_suboptimal:2020} to model expected annual damages for a particular house.

\subsubsection{Multiple PDFs}

\subsubsection{Candidate decisions}

\subsubsection{States of the World (SOWs)}

\subsubsection{System model}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{historic_surge.pdf}
    \caption{
        Time series of annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
    }
    \label{fig:observations}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=6in]{hazard-tradeoffs.pdf}
    \caption{
        Neglecting hydrodynamics, we add sea level rise (a) to a stationary Bayesian GEV model of storm surge (b) to get flood hazard.
        The convolution of hazard and fragility (c) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
    }
\end{figure}

\subsection{One size does not fit all}

\begin{itemize}
    \item Tailoring guidance to specific building characteristics can improve outcomes
    \item Elevating to ``\acrshort{bfe} plus a foot'' is not always optimal \citep{xian_elevation:2017,zarekarizi_suboptimal:2020}
    \item Both over- and under-building can be costly \citep{ansar_bigisfragile:2017,DossGollin:2019}
\end{itemize}
\begin{figure}
    \centering
    \begin{subfigure}{0.3\textwidth}
        \subcaption{\SI{2.5}{ft} below BFE}
        \includegraphics[width=\textwidth]{5.5.pdf}
    \end{subfigure}~
    \begin{subfigure}{0.3\textwidth}
        \subcaption{\SI{0.5}{ft} below BFE}
        \includegraphics[width=\textwidth]{7.5.pdf}
    \end{subfigure}
    \caption{
        Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \citep{wong_brick0.2:2017} dynamics).
        Note unequal $y$-axis scales.
    }
\end{figure}

\subsection{The multiple PDF problem}

Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie, the \textbf{\itshape{Multiple PDF Problem}}.
Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
\begin{figure}
    \centering
    \includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:boxplots}
\end{figure}

\subsection{It matters which PDF you choose}

Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes.
\begin{figure}
    \centering
    \includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
    \includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
    \caption{
        Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
    }
    \label{fig:tradeoffs}
\end{figure}
Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement.

\section{Synthesizing PDFs for decision relevance}\label{sec:prior-averaging}

\subsection{Methods}

While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability \citep[see][]{schneider_dangerous:2001,schneider_scenarios:2002}.
\begin{figure}
    \centering
    \includegraphics[width=4in]{bayes-rdm.pdf}
    \caption{
        ``Prior Model Averaging'' framework.
    }\label{fig:flowchart}
\end{figure}
Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
A system model $(f)$ quantifies the performance $(u)$ of a decision ($x$) under a single SOW ($s$).
To synthesize, we weight PDFs using an approach based on Bayesian Model Averaging \citep{wong_surge:2018,Yao:2018bu}.

\subsection{Findings}

\begin{figure}
    \centering
    \includegraphics[width=6in]{inference_weights.pdf}
    \caption{
        The weights assigned to each PDF (from \cref{fig:boxplots}) under each prior considered.
    }
\end{figure}
``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.

\section{Conclusions}\label{sec:conclusions}

To summarize:
\begin{itemize}
    \item Natural extensions to Bayesian updating
    \item Belief $p(s)$ can't be ``right''  \citep{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \citep{quinn_exploratory:2020}
    \item Instead, we can ask \textbf{what assumptions are different priors consistent with?}
\end{itemize}

\subsection{Implications for policy}

\begin{figure}
    \centering
    \includegraphics[width=3in]{lsl_priors.pdf}~
    \includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
    \caption{
        The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
        Alternative priors can provide robustness checks.
    }
\end{figure}

\subsection{Future needs}
This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
\textbf{\scshape let's collaborate on:}
\begin{itemize}
    \item More complex models to capture more relevant metrics
    \item Better models of nonstationary hazards
    \item Interacting, sequential decisions
\end{itemize}

@@

\begin{figure}
    \centering
    %\includegraphics[height=2in]{uconn-elevateA}
    \caption{
        Conceptual sketch illustrating the path from historical data to FEMA's guidance.
        This figure \emph{needs to be revised} (KATERINA?) to have three panels, of which this is the third.
        (a): A time series of historical floods.
        (b): A \gls{pdf} of the flood peak.
        (c): as shown, but with waves eliminated.
        This sketch figure is from \citet{patten_elevate:2016}.
    }\label{fig:house-sketch}
\end{figure}

@@

\begin{figure}
    \centering
    %\includegraphics[width=\textwidth, height=4.5in, keepaspectratio=true]{conceptual.pdf}
    \caption{
        Schematic of our decision support framework.
        Nodes in green are modeling choices that the modeler must make while those in yellow are computed quantities.
        Teal text illustrates a non-exhaustive set of opportunities to perform model checks and/or engage with stakeholders to iterate on design of the decision support system.
        Feedbacks between these questions and the green nodes are omitted for clarity.
        Our approach could be extended by coupling the evaluation framework to an optimization or policy search tool that updates the candidate decisions.
        This approach could also be extended to a stochastic system model.
    }\label{fig:framework}
\end{figure}

@@

\begin{figure}
    \centering
    %\includegraphics[height=2.5in, width=0.6\textwidth, keepaspectratio=true]{xlrm}
    \caption{
        Diagram showing uncertainties ($X$; red), levers ($L$; yellow), system relationships ($R$; gray), and metrics ($M$; blue).
    }\label{fig:xlrm}
\end{figure}
